\LARGE \bf Enhancing Security in Multi-Robot Systems through Co-Observation Planning, Reachability Analysis, and Network Flow

### 2024-12-09 response.tex

%% Sets page size and margins
\usepackage[a4paper,top=2.5cm,bottom=2cm,left=1.7cm,right=1.7cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
% \usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{indentfirst}
\usepackage{graphicx}
% \usepackage{subfigure}
\usepackage{float}
\usepackage{threeparttable} 
\usepackage{multirow}
\usepackage{url}
\usepackage{xcolor}
\pagenumbering{arabic}

\usepackage[most]{tcolorbox}
\newcommand{\rrtstar}{$\texttt{RRT}^\texttt{*}$}
\newcommand{\SubItem}[1]{
    {\setlength\itemindent{15pt} \item[-] #1}
}

\newtheorem{constraint}{ADMM constraint}
\newtheorem{example}{Example}
\newcommand{\re}{\tcblower \underline{\textbf{Response:}}}
\newcommand{\rv}{{\large{\underline{\textbf{Revision:}}}}\quad}
\usepackage[export]{adjustbox} 
\usepackage[capitalise]{cleveref}
\usepackage{xcolor}
\newcommand{\new}[1]{\textcolor{blue}{#1}}
\newcommand{\news}{\color{blue}}
\graphicspath{{Figs/}}

@@

\begin{cmt}{}{}
This reachability ellipsoid is an over approximation of the exact
reachability region in Definition 3.”; A proof or at least a
reference to the proof should be provided.
\re While we believe that a formal proof is not necessary for this context, we have added additional content to make the approximation more intuitive. This includes a clearer explanation of how the reachability ellipsoid relates to the exact reachability region. These additions aim to enhance the reader's understanding without requiring formal proof.
\end{cmt}
\rv {\news
\begin{definition}\label{sec:ellipsoidal definition}
Consider a robot $i$ starting from $q_{1}$ at time $t_1$ and reaching $q_{2}$ at time $t_2$. The \emph{reachability region} between $t_1$ and $t_2$ is defined as the set of points $q'$ in the free configuration space for which there exists a kinematically feasible trajectory containing $q'$.
\end{definition}

For simplicity, in this paper we consider only a maximum velocity constraint $v_{max}$; the reachability region for $q'$ can then be over-approximated by the following (the over-approximation is obtained by neglecting physical obstacles):
\begin{definition}\label{def:Reachability}
	The \emph{reachability ellipsoid $\cE$} is defined as the region $\mathcal{E}(q_1,q_2,t_{1},t_{2})=\{\tilde{q}\in\mathbb{R}^n: d(q_1,\tilde{q})+d(\tilde{q},q_2)<2a\}$, where $a=\frac{v_{max}}{2}(t_2-t_1)$.
\end{definition}
This region is an ellipsoid because it represents the set of points $q'$ whose sum of distances two \emph{foci} $q_1$ and $q_2$ is less than $2a$, where a is the major radius of the ellipsoid. }
\begin{cmt}{}{}
	Other grammar related comments.
	\re We have carefully reviewed and corrected all listed grammar mistakes, as well as others that were not explicitly mentioned. 
\end{cmt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Rewiewer 2
\newpage
\section{Response to reviewer 2}

%%%%%%%%%%%%% Comment 2.1
\begin{cmt*}{}{}
	This paper proposed a novel multi-robot optimal planning algorithm that
	integrated mutual observations and introduced reachability constraints
	for enhanced security, which ensures that, even with adversarial
	movements, compromised robots cannot breach forbidden regions without
	missing scheduled co-observations. The contribution of this paper is
	hard evaluate due to the organization. 
	\re We sincerely thank the reviewer for their valuable comments and feedback. We acknowledge the concern regarding the organization of the manuscript, which may have made it challenging to fully evaluate the contributions. In the revised version, we have carefully restructured the manuscript to improve its organization and readability, ensuring that the key contributions are presented more clearly and effectively. We greatly appreciate your valuable feedback, which has been instrumental in enhancing the quality of our work.
\end{cmt*}
%%%%%%%%%%%%% Comment 2.1
\begin{cmt}{}{}
The dimension of p is erroneously defined;
\re After carefully reviewing the manuscript, we believe the reviewer might have meant the dimension of $q$ rather than $p$. In the revised version, we have corrected the definition of $q$ to ensure clarity and accuracy. Thank you for bringing this to our attention.
\end{cmt}
\rv {\news We denote the trajectory as $\vq_i = [q_{i0}\dots q_{iT}]\in\real{m\times T}$, where $q_{ij}\in\real{m}$ is the waypoint of agent $i$ in a $m$ dimensional state space, $T$ is the time horizon. For a total of $n_p$ robots, trajectories can be represented as an aggregated vector $\vq = \stack(\vq_1,\dots,\vq_{n_p})\in \real{m n_p\times T}$, where $\stack(\cdot)$ denotes the vertical stacking operation. }

%%%%%%%%%%%%% Comment 2.3
\begin{cmt}{}{}
There are many grammar errors, please check it carefully
\re We have carefully reviewed and corrected all identified grammar issues, as well as others not explicitly mentioned. The revised manuscript has undergone thorough proofreading to ensure clarity and grammatical accuracy.

%%%%%%%%%%%%% Comment 2.4
\begin{cmt}{}{}
In Definition 1, is reasonable to say ``any potential deviations''?
\re We believe that this phrase is resonable and essential to the definition and intention of security against plan-deviation attacks. The definition is designed to ensure that the trajectory plan accounts for all possible paths a compromised robot might take toward forbidden regions. By explicitly requiring that any such deviation results in the robot missing its next co-observation, the definition provides a rigorous and comprehensive security guarantee. As such, the phrase "any potential deviations" accurately reflects the intended scope of the definition.
\end{cmt}

\begin{cmt}{}{}
	In the definition of $U$, what does ``,'' mean?
	\re  We are not entirely sure what the specific concern or issue is in this context. However, if the reviewer is referring to $u'$ , it was intended to denote the vector $u$ before normalization. To avoid any potential confusion, we have revised the notation in the updated manuscript. We appreciate your feedback and have taken steps to ensure clarity in this definition.
\end{cmt}
\setcounter{definition}{8}
\setcounter{equation}{41}
\rv {\news \begin{definition} Let $\nu_1$ and $\nu_2$ be two unitary vectors ($\norm{\nu_1}=\norm{\nu_2}=1$). Define the normalized vector $u = \frac{\nu_1+\nu_2}{\norm{\nu_1+\nu_2}}$,
	% \begin{equation}
	%   u=\nu_1+\nu_2,\quad u =\frac{u'}{\norm{u'}}.
	% \end{equation}
  the \emph{Householder rotation} $H(\nu_1,\nu_2)$ is defined as
	\begin{equation}\label{eq:H definition}
	  H(\nu_1,\nu_2) = 2 u u\transpose-I.
	\end{equation}
  \end{definition}}
\begin{cmt}{}{}
	How to solve (9a)?
	\re  In practical applications, Eq.4a (originally 9a) is solved iteratively using a nonlinear optimization solver, such as \emph{fmincon}. Additionally, the derivatives of the constraint projection functions, as derived later in the manuscript, are specifically utilized to solve the optimization problem more efficiently and quickly. These derivatives enhance the solver's performance by providing necessary gradient information, improving convergence speed. We have clarified these points in the revised manuscript to provide further insight into the solution process.
\end{cmt}
\rv {\news
\setcounter{remark}{0}
\begin{remark}
In practical application, (4a) is solved iteratively via a nonlinear optimization solver like \emph{fmincon}. The gradients of the constraint functions, derived later, is used to enhance solver efficiency and accuracy. Meanwhile, (4b) and (4c) have closed form solutions.
\end{remark}
}
\begin{cmt}{}{}
	How to define $d_{\max}$ in (14)?
	\re $d_{max}$ is the maximum distance between a pair of robots that the secure can be ensured through co-observation.
\end{cmt}
\setcounter{constraint}{0}
\rv \begin{constraint}[Co-observation constraint]\label{constraint:coobservation}
	\begin{align}
	D(\vq) &= \overrightarrow{q_{aj}q_{bj}}, \label{eq:coobservation_constraint}\\
	  \sZ &= \{\vz \mid \norm{\vz} \leq d_{max} \},\\
	   \Pi_\sZ(z) & = \begin{cases}
	d_{max}\frac{\vz}{\norm{\vz}} &\text{if } \norm{\vz} > d_{max},\\
	\vz	& otherwise,
	\end{cases}
	\end{align}
	where $a,b$ are the indices of the pair of agents required for a mutual inspection, \new{$d_{max}$ is the maximum distance between a pair of robots that the secure can be ensured through co-observation.}
	\end{constraint}

\begin{cmt}{}{}
	In Definition 4, it is unclear that $\mathcal E$ is an ellipsoid. There is no definition of the function $d(\cdot,\cdot)$;
	\re In the revised manuscript, we have added additional explanations to clarify that $\cE$ is an ellipsoid and to explicitly define the function $d(\cdot,\cdot)$ to avoid confusion and ensure clarity.
\end{cmt}
\rv {\news \setcounter{definition}{1}
\begin{definition}\label{sec:ellipsoidal definition}
Consider a robot $i$ starting from $q_{1}$ at time $t_1$ and reaching $q_{2}$ at time $t_2$. The \emph{reachability region} between $t_1$ and $t_2$ is defined as the set of points $q'$ in the free configuration space for which there exists a kinematically feasible trajectory containing $q'$.
\end{definition}

For simplicity, in this paper we consider only a maximum velocity constraint $v_{max}$; the reachability region for $q'$ can then be over-approximated by the following (the over-approximation is obtained by neglecting physical obstacles):
\begin{definition}\label{def:Reachability}
	The \emph{reachability ellipsoid $\cE$} is defined as the region $\mathcal{E}(q_1,q_2,t_{1},t_{2})=\{\tilde{q}\in\mathbb{R}^n: d(q_1,\tilde{q})+d(\tilde{q},q_2)<2a\}$, where $a=\frac{v_{max}}{2}(t_2-t_1)$, and $d(\cdot,\cdot)$ denotes the Euclidean distance between two points.
\end{definition}
This region is an ellipsoid because it represents the set of points $q'$ whose sum of distances two \emph{foci} $q_1$ and $q_2$ is less than $2a$, where a is the major radius of the ellipsoid. }
\end{document}

### 2024-12-09 paper.tex
update

The update steps of the algorithm are then derived as \cite{Boyd2011}:
\begin{subequations}\label{eq:ADMMupdate}
	\begin{align}
		\vq^{k+1}&:=\argmin_\vq(\varPhi(\vq^k) +\frac{\rho}{2}\norm{D(\vq)-\vz^k+\vu^k}_2^2),\label{eq:admm-mod-update-q}\\
		\vz^{k+1}&:=\Pi_\mathcal{Z}(D(\vq^{k+1})+\vu^k), \label{eq:z-update} \\
		\vu^{k+1}&:=\vu^k+D(\vq^{k+1})-\vz^{k+1}, \label{eq:u-update}
	\end{align}
\end{subequations}
where $\Pi_\mathcal{Z}$ is the new projection function to the modified constraint set $\mathcal{Z}$, $\vu$ represents a scaled dual variable that, intuitively, accumulates the sum of primal residuals
\begin{equation}\label{eq:primal-residual}
	\vr^{k}=D(\vq^{k+1})-\vz^{k+1}.
\end{equation}
Checking the primal residuals alongside the dual residuals 
\begin{equation}\label{eq:dual-residual}
	\vs^{k}=-\rho(\vz^{k}-\vz^{k-1})
\end{equation}
after each iteration, the steps are reiterated until convergence when the primal and dual residuals are small, or divergence when primal and dual residuals remain large after a fixed number of iterations.
{\news
\begin{remark}
In practical application, \eqref{eq:admm-mod-update-q} is solved iteratively via a nonlinear optimization solver like \emph{fmincon}, while \eqref{eq:z-update} and~\eqref{eq:u-update} have closed form solutions.
\end{remark}
}
% \rtron{Maybe state here that (9a) is solved iteratively, while (9b) and (9c) are closed-form.}

@@

For simplicity, in this paper we consider only a maximum velocity constraint $v_{max}$; the reachability region for $q'$ can then be over-approximated by the following (the over-approximation is obtained by neglecting physical obstacles):%$\mathcal{E}=\{q'\in\mathbb{R}^n: d(q_1,q')+d(q'q_2)<v_{max}(t_2-t_1)\}$, where $d(\cdot,\cdot)$ denotes the Euclidean distance between two points.
\begin{definition}\label{def:Reachability}
	The \emph{reachability ellipsoid $\cE$} is defined as the region $\mathcal{E}(q_1,q_2,t_{1},t_{2})=\{\tilde{q}\in\mathbb{R}^n: d(q_1,\tilde{q})+d(\tilde{q},q_2)<2a\}$, where $a=\frac{v_{max}}{2}(t_2-t_1)$.
\end{definition}
This region is an ellipsoid because it represents the set of points $q'$ whose sum of distances two \emph{foci} $q_1$ and $q_2$ is less than $2a$, where a is the major radius of the ellipsoid. }
 \begin{figure}
    \centering
    \subfloat[Showcase of a reachability region\label{fig:EllipseConstraintExample}]{\includegraphics[width=0.47\linewidth, trim = 2cm 2.5cm 2cm 5cm]{Reachability}}
    \subfloat[Point-ellipsoid constraint \label{fig:Ellipse-to-point}]{\includegraphics[width=0.47\linewidth, trim = 0.3cm 1.3cm 0.3cm 1.3cm, clip]{Ellipse2point}}

@@

{\news The condition that a reachability ellipsoid $\cE(q_1,q_2)$ does not intersect with any forbidden region is sufficient to secure the trajectory of the robot between $q_1$ and $q_2$ according to \cref{def:secured-plan}: any deviation to a point $p_o$ outside $\cE(q_1,q_2)$ will cause the robot to miss a potential observation at $p_2$, $t_2$. \cref{def:secured-plan} can then be restated as follows:
\begin{definition}\label{rmk:revised-security}
  A multi-robot trajectory is secured against plan-deviation attacks if there exists a co-observation plan such that the reachability region between each consecutive co-observation does not intersect with any forbidden regions.
\end{definition}
}

\subsection{Transformation to canonical coordinates}\label{sec:rotation2Standard}
\newcommand{\oFE}{o}
{\news To use the definition of reachability ellipsoid from the section above as computational constraints for ADMM, we first apply a differentiable rigid body transformation to reposition the ellipse $\cE$ from the global frame $\cF$ to a canonical frame $\cF_\cE$, where the origin of $\cF_\cE$ is at the ellipsoid's center $\oFE = \frac{1}{2}(q_1+q_2)$, and the first axis of $\cF_\cE$ is aligned with the foci (see \cref{fig:EllipseConstraintExample} for an illustration). 
From this definition, a unit vector along the first axis of the ellipsoid in the frames $\cF$ and $\cF_\cE$ is given by $\nu_\cF = \frac{q_2-q_1}{\norm{q_2-q_1}}$ and $\nu_\cE =[1,0,0]^T$, respectively.
The full transformation from coordinates $q^\cE\in\cF$ to $q\in\cF$, and its inverse, are then parametrized by a rotation $R^\cF_\cE$ and a translation $o^\cF_\cE$ as (we drop the subscript and superscript from $R^\cF_\cE$ and $o^\cF_\cE$ to simplify the notation):
\begin{equation}\label{eq:transformations}
  q=R\Eframe{q}+\oFE,\quad
  \Eframe{q}=R\transpose(q-\oFE).
\end{equation}
where the rotation matrix $R=H(\nu_\cF(q_1,q_2),\nu_\cE)$ is a \emph{Householder rotation} $H$ (a differentiable linear transformation describing the minimal rotation between the two vectors, see \cref{sec:householder} for details).  To simplify the notation, in the following, we will consider $H$ to be a function of $q_1,q_2$ directly, i.e. $H(q_1,q_2)$. More details on \eqref{eq:transformations} are given in \cref{apx:transformation}.
  % We define $\nu_\cF$ and $\nu_\cE$ to represent the $x$-axis unitary vector of $\cF_\cE$ in the frames $\cF$ and $\cF_\cE$, respectively. Formally:
  % \begin{align}
  %   \nu_\cF &= \frac{q_2-q_1}{\norm{q_2-q_1}}, & \nu_\cE&=[1,0,0]^T;
  % \end{align}
  % \new{see Fig.\ref{fig:EllipseConstraintExample} for an illustration. Note that $\nu_{\cF}$ is a function of $q_1,q_2$ while $\nu_\cE$ is constant. From the definition of $\cF_\cE$, the rotation $R$ can be found using a \emph{Householder rotation} (see \cref{sec:householder}), while from \eqref{eq:transformations} the translation $o$ must be equal to the center of $\cE(q_1,q_2)$ expressed in $\cF$, i.e.:}
  % \begin{align} 
  %   R&=H(\nu_\cF(q_1,q_2),\nu_\cE), &o&=\frac{1}{2}(q_1+q_2).
  % \end{align}

>>>>

The update steps of the algorithm are then derived as \cite{Boyd2011}:
\begin{subequations}\label{eq:ADMMupdate}
	\begin{align}
		\vq^{k+1}&:=\argmin_\vq(\varPhi(\vq^k) +\frac{\rho}{2}\norm{D(\vq)-\vz^k+\vu^k}_2^2),\label{eq:admm-mod-update-q}\\
		\vz^{k+1}&:=\Pi_\mathcal{Z}(D(\vq^{k+1})+\vu^k), \label{eq:z-update} \\
		\vu^{k+1}&:=\vu^k+D(\vq^{k+1})-\vz^{k+1}, \label{eq:u-update}
	\end{align}
\end{subequations}
where $\Pi_\mathcal{Z}$ is the new projection function to the modified constraint set $\mathcal{Z}$, $\vu$ represents a scaled dual variable that, intuitively, accumulates the sum of primal residuals
\begin{equation}\label{eq:primal-residual}
	\vr^{k}=D(\vq^{k+1})-\vz^{k+1}.
\end{equation}
Checking the primal residuals alongside the dual residuals 
\begin{equation}\label{eq:dual-residual}
	\vs^{k}=-\rho(\vz^{k}-\vz^{k-1})
\end{equation}
after each iteration, the steps are reiterated until convergence when the primal and dual residuals are small, or divergence when primal and dual residuals remain large after a fixed number of iterations.
{\news
\begin{remark}
In practical application, \eqref{eq:admm-mod-update-q} is solved iteratively via a nonlinear optimization solver like \emph{fmincon}. The gradients of the constraint functions, derived later, is used to enhance solver efficiency and accuracy. Meanwhile, \eqref{eq:z-update} and~\eqref{eq:u-update} have closed form solutions.
\end{remark}
}
% \rtron{Maybe state here that (9a) is solved iteratively, while (9b) and (9c) are closed-form.}

@@

For simplicity, in this paper we consider only a maximum velocity constraint $v_{max}$; the reachability region for $q'$ can then be over-approximated by the following (the over-approximation is obtained by neglecting physical obstacles):%$\mathcal{E}=\{q'\in\mathbb{R}^n: d(q_1,q')+d(q'q_2)<v_{max}(t_2-t_1)\}$, where $d(\cdot,\cdot)$ denotes the Euclidean distance between two points.
\begin{definition}\label{def:Reachability}
	The \emph{reachability ellipsoid $\cE$} is defined as the region $\mathcal{E}(q_1,q_2,t_{1},t_{2})=\{\tilde{q}\in\mathbb{R}^n: d(q_1,\tilde{q})+d(\tilde{q},q_2)<2a\}$, where $a=\frac{v_{max}}{2}(t_2-t_1)$, and $d(\cdot,\cdot)$ denotes the Euclidean distance between two points.
\end{definition}
This region is an ellipsoid because it represents the set of points $q'$ whose sum of distances two \emph{foci} $q_1$ and $q_2$ is less than $2a$, where a is the major radius of the ellipsoid. }
 \begin{figure}
    \centering
    \subfloat[Showcase of a reachability region\label{fig:EllipseConstraintExample}]{\includegraphics[width=0.47\linewidth, trim = 2cm 2.5cm 2cm 5cm]{Reachability}}
    \subfloat[Point-ellipsoid constraint \label{fig:Ellipse-to-point}]{\includegraphics[width=0.47\linewidth, trim = 0.3cm 1.3cm 0.3cm 1.3cm, clip]{Ellipse2point}}

@@

{\news The condition that a reachability ellipsoid $\cE(q_1,q_2)$ does not intersect with any forbidden region is sufficient to secure the trajectory of the robot between $q_1$ and $q_2$ according to \cref{def:secured-plan}: any deviation to a point $p_o$ outside $\cE(q_1,q_2)$ will cause the robot to miss a potential observation at $p_2$, $t_2$. \cref{def:secured-plan} can then be restated as follows:
\begin{definition}\label{rmk:revised-security}
  A multi-robot trajectory is secured against plan-deviation attacks if there exists a co-observation plan such that the reachability region between each consecutive co-observation does not intersect with any forbidden regions.
\end{definition}
}
\subsection{Transformation to canonical coordinates}\label{sec:rotation2Standard}
\newcommand{\oFE}{o}
{\news To use the definition of reachability ellipsoid from the section above as computational constraints for ADMM, we first apply a differentiable rigid body transformation to reposition the ellipse $\cE$ from the global frame $\cF$ to a canonical frame $\cF_\cE$, where the origin of $\cF_\cE$ is at the ellipsoid's center $\oFE = \frac{1}{2}(q_1+q_2)$, and the first axis of $\cF_\cE$ is aligned with the foci (see \cref{fig:EllipseConstraintExample} for an illustration). 
From this definition, a unit vector along the first axis of the ellipsoid in the frames $\cF$ and $\cF_\cE$ is given by $\nu_\cF = \frac{q_2-q_1}{\norm{q_2-q_1}}$ and $\nu_\cE =[1,0,0]^T$, respectively.
The full transformation from coordinates $q^\cE\in\cF$ to $q\in\cF$, and its inverse, are then parametrized by a rotation $R^\cF_\cE$ and a translation $o^\cF_\cE$ as (we drop the subscript and superscript from $R^\cF_\cE$ and $o^\cF_\cE$ to simplify the notation):
\begin{equation}\label{eq:transformations}
  q=R\Eframe{q}+\oFE,\quad
  \Eframe{q}=R\transpose(q-\oFE).
\end{equation}
where the rotation matrix $R=H(\nu_\cF(q_1,q_2),\nu_\cE)$ is a \emph{Householder rotation} $H$ (a differentiable linear transformation describing the minimal rotation between the two vectors, see \cref{sec:householder} for details).  To simplify the notation, in the following, we will consider $H$ to be a function of $q_1,q_2$ directly, i.e. $H(q_1,q_2)$. More details on \eqref{eq:transformations} are given in \cref{apx:transformation}.
  % We define $\nu_\cF$ and $\nu_\cE$ to represent the $x$-axis unitary vector of $\cF_\cE$ in the frames $\cF$ and $\cF_\cE$, respectively. Formally:
  % \begin{align}
  %   \nu_\cF &= \frac{q_2-q_1}{\norm{q_2-q_1}}, & \nu_\cE&=[1,0,0]^T;
  % \end{align}
  % \new{see Fig.\ref{fig:EllipseConstraintExample} for an illustration. Note that $\nu_{\cF}$ is a function of $q_1,q_2$ while $\nu_\cE$ is constant. From the definition of $\cF_\cE$, the rotation $R$ can be found using a \emph{Householder rotation} (see \cref{sec:householder}), while from \eqref{eq:transformations} the translation $o$ must be equal to the center of $\cE(q_1,q_2)$ expressed in $\cF$, i.e.:}
  % \begin{align} 
  %   R&=H(\nu_\cF(q_1,q_2),\nu_\cE), &o&=\frac{1}{2}(q_1+q_2).
  % \end{align}
